---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gtfssupplyindex

<!-- badges: start -->
<!-- badges: end -->

The goal of gtfssupplyindex is to provide tools to calculate 
the Transit Supply Index as per 
Graham Currie (2010) 
Quantifying spatial gaps in public transport supply based on social needs. 
Journal of Transport Geography, 18 (1):31–41, 2010. ISSN 0966-6923. 
doi: https://doi.org/10.1016/j.jtrangeo.2008.12.002. 
URL https://www.sciencedirect. com/science/article/pii/S0966692308001518.


## Installation

You can install the development version of gtfssupplyindex from [GitHub](https://github.com/) with:



``` r
# install.packages("devtools")
devtools::install_github("James-Reynolds/gtfssupplyindex")
```

## Example

You can calculate the SI by day, hour, route and area of interest for 
the Mornington Peninsula Tourist Railway (ncluded as a sample GTFS feed) 
as follows:

```{r example}
library(gtfssupplyindex)
####-----first load all the inputs
#load the revised mornington GTFS data
list_gtfs = gtfssupplyindex:::gtfs_by_route_type(system.file("extdata/mornington180109",
"gtfs.zip", 
package = "gtfssupplyindex", 
mustWork = TRUE))
 
areas_of_interest <- load_areas_of_interest(absmapsdata::sa22021 %>% 
           dplyr::filter(sa3_name_2021 ==  "Mornington Peninsula") %>% 
                                              dplyr::select(sa2_code_2021),  
                                            area_id_field = "sa2_code_2021") 
 buffer_distance <- gtfssupplyindex:::load_buffer_zones()
 
 stops_in_or_near_areas <- gtfssupplyindex:::stops_in_walk_dist(
 list_gtfs = list_gtfs, 
 areas_of_interest = areas_of_interest,
 EPSG_for_transform = 28355,
 verbose = FALSE
 )
 
 SI_by_day_hour_and_route_type(
 list_gtfs, 
 stops_in_or_near_areas, 
 "2018-12-23",
 "2018-12-30",
 verbose = FALSE)

```

See also the vignette at https://github.com/James-Reynolds/gtfssupplyindex/tree/main/vignettes
