
---
title: "Accessibility by and to transit; comparison with transit Supply Index results"
runningheader: "Reynolds (2025)" # only for pdf output
author: "James Reynolds"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
bibliography: [packages.bib, References.bib]
link-citations: yes
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=2in,height=2in]{ptrg-logo-s.png}\LARGE\\}
  - \posttitle{\end{center}}

---

```{r setup, include=FALSE}
library(tufte)

library(tidyverse)
library(tidytransit)
library(sp)
library(strayr)
library(ptinpoly)
library(magrittr)
library(ggplot2)
library(sf)
library(ASGS.foyer)
library(raster)
library(ggmap)
library(units)
library(janitor)
library(mapview)
library(ggstatsplot)
library(gtsummary)
library(moments)
library(scales)
library(gtfstools)
library(lubridate)
library(kableExtra)
library(knitr)
library(readxl)
library(dplyr)
library(devtools)
load_all()
# invalidate cache when the tufte version changes
#knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
```

# Introduction
@currie2007identifying and @currie2010identifying previously developed a 
transit Supply Index (SI) to compare public transport frequency and coverage between 
different geographic areas. In subsequent work, @Reynolds:2025aa  
developed code to calculate SI scores 
from GTFS datasets 
as an R package [@gtfssupplyindex_github]. 
SI scores for all of Victoria for the week starting 27/04/23 were also 
provided as an input to a forthcoming paper on how transit service levels impact housing prices[@liang5144567exploring]. 


The rest of this document 
is structured as follows: 


# Research context


## The Suppy Index

```{marginfigure}
\begin{equation}
  SI_{area, time} = \sum{\frac{Area_{Bn}}{Area_{area}}*SL_{n, time}}
  \end{equation}
```

The Supply Index (SI) equation 
is shown in the margin figure^[
Minor adjustments have been made 
to generalise the equation, 
as Currie and Senbergs (2007) focus 
was the context of Melbourne's Census Collection Districts (CCD) 
and calculations based on a week of transit service. 
CCDs predate the introduction of 
Statistical Areas 1, 2, 3, and 4 (SA1, SA2, SA3, SA4), 
and other geographical divisions 
currently used by the Australian Bureau of Statistics (ABS), 
which may be more familiar to readers.], 
in which:
(1) $SI_{area, time}$ is the Supply Index for the area of interest 
and a given period of time;
(2) $Area_{Bn}$ is the buffer area for each stop (n) within the area of interest. 
In Currie and Senbergs (2007) this was based on 
a radius of 400 metres for bus and tram stops, 
and 800 metres for railway stations;
(3) $Area_{area}$ is the area of the area of interest; and
(4) $SL_{n,time}$ is the number of transit arrivals for each stop 
for a given time period.


# Methodology
R [@R-base], a widely used and readily available  statistical programming language, was adopted for code development. The package development setup and workflow described by @wickham2023r 
was adopted in this study. 

Various existing packages were relied upon including: the sf package [@R-sf] for geospatial analysis; the tidyverse [@tidyverse2019]; gtfstools [@R-gtfstools]; and tidytransit [@R-tidytransit]. Some code was adapted from the tidytransit, gtfstools and other package's examples, vignettes and other documentation.
Tidytransit is licensed under GPL-2 | GPL-3, 
while gtfstools is lincensed under the more permissive MIT license. 
For simplicity, 
the gtfssupplyindex package developed here 
is licensed under GPL-3 
so as to match tidytransit. 

## Cases

```{r mornington_map, fig.margin = TRUE, fig.cap = "Mornington Penninsula SA1 zones and location of Mornington Tourist Rail stops.", echo = FALSE, warning=FALSE, message=FALSE, cache=TRUE}
### ---------------- get abs data for Mornington Peninsula
#options(timeout = 1000)
#remotes::install_github("wfmackey/absmapsdata")

#get_mornington_sa1 <- function(){
#  mornington_sa12021 <- absmapsdata::sa12021 %>% filter(sa3_name_2021 == "Mornington Peninsula") %>% select(sa1_code_2021)
#  sf::st_write(mornington_sa12021, "inst/extdata/mornington_sa12021.geojson", append=FALSE)
#  return(mornington_sa12021)
#}
#get_mornington_sa3 <- function(){
#  mornington_sa32021 <- absmapsdata::sa32021 %>% filter(sa3_name_2021 == "Mornington Peninsula") %>% select(sa3_code_2021)
#  sf::st_write(mornington_sa32021, "inst/extdata/mornington_sa32021.geojson", append=FALSE)
#  return(mornington_sa32021)
#}

### ---------------- load SA3 abs maps data for just mornington peninsula
areas_of_interest <- load_areas_of_interest(areas_of_interest = absmapsdata::sa12021 %>% filter(sa3_name_2021 == "Mornington Peninsula") %>% select(sa1_code_2021), 
  area_id_field = "sa1_code_2021")

# map the areas_of_interest
map <- areas_of_interest %>% 
  ggplot() +
  geom_sf(aes(geometry = geometry))
#map
#set the EPSG to transform from lat/lon to metres
EPSG_for_transform = 28355

### ---------------------------- load and update the route_types in the mornington gtfs data
mornington_2019 <- tidytransit::read_gtfs("inst/extdata/mornington180109/gtfs.zip")
#deal with issue to do with missing route_color information
mornington_2019$routes$route_color <- "Blue"
#convert route type from 107 (tourist railway) to 2 (rail) ---NEW FEATURE NEEDED --- EXPAND HANDLING OF MODES TO ENABLE CALCULATION OF MODES FROM EXTENDED GTFS LIST
mornington_2019$routes$route_type <- as.integer(2)
# write back to zip file
tidytransit::write_gtfs(mornington_2019, "inst/extdata/mornington180109/gtfs.zip")

#load the revised mornington GTFS data
list_gtfs = gtfssupplyindex:::gtfs_by_route_type("inst/extdata/mornington180109/gtfs.zip")

stops_as_sf_mornington <-  list_gtfs[[1]]$stops %>% 
  tidytransit::stops_as_sf()

#map the stops on the ABS data
map + 
  geom_sf(data = stops_as_sf_mornington, aes(geometry = geometry))

stops_in_or_near_areas <- gtfssupplyindex:::stops_in_walk_dist(
  list_gtfs = list_gtfs, 
  areas_of_interest = areas_of_interest,
  EPSG_for_transform = 28355
)

```

Code development was undertaken using a number of cases for testing purposes, being: the  Mornington Penninsula Tourist Railway, and Melbourne. 


The Morning Penninsula Tourist Railway runs on Sundays and Wednesdays between Moorooduc and Mornington, with an intermediate stop at Tanti Park^[https://transitfeeds.com/p/mornington-railway/806/latest/stops]. A GTFS feed from 2018 was selected. ABS data was also used, 
primarily through the strayr 
and absmapsdata packages [@r-strayr].
The Mornington Peninsular SA3 zone and the SA1 zones contained within were adopted as the areas_of_interest. 

Larger scale testing was performed using the Victorian GTFS feed, 
published by Public Transport Victoria (PTV), 
sourced via @transitfeeds_victoria:2023aa for historical feeds. 
Again, ABS data was used for the areas_of_interest. 



# Results

# Discussion and conclusions




# References {#references }


```{r, include=FALSE}
knitr::write_bib(file = 'packages.bib')
```


